\chapter{$U(1)_{T^3_R}$ Gauge Extension of the Standard Model}\label{ch:U1T3R}

Extensions of the SM that introduce new $U(1)$ gauge symmetries are among the most widely studied scenarios to address phenomenological hints while maintaining theoretical consistency. Some natural candidates to extend the SM gauge group via a $U(1)$ factor arise from the fact that the SM exhibits several global symmetries that are not imposed by its gauge structure but emerge naturally from the renormalizable interactions allowed by the gauge group $SU(3)_C \times SU(2)_L \times U(1)_Y$. Among these, baryon number ($B$) and lepton number ($L$) conservation are particularly notable. In the SM Lagrangian, these quantum numbers are conserved at the perturbative level, making them \textit{accidental symmetries}. However, they are not protected by any fundamental principle~\parencite{Weinberg:1979sa, Wilczek:1979hc}.

The accidental nature of these symmetries motivates their promotion to local gauge symmetries, which would forbid dangerous operators that violate $B$ and $L$ at the renormalizable level, offer mechanisms for generating small neutrino masses, and address cosmological puzzles such as the matter-antimatter asymmetry~\parencite{Marshak:1979fm, Mohapatra:1980yp, Buchmuller:2004nz}. However, not all combinations of $B$ and $L$ can be consistently gauged. A fundamental requirement for any gauge theory is the absence of gauge anomalies, which arise from triangle diagrams with chiral fermions and would render the quantum theory inconsistent~\parencite{Adler:1969gk, Bell:1969ts, tHooft:1980xss}. %For a $U(1)$ gauge symmetry with charge $Q$, anomaly cancellation requires $\sum_f Q_f^3 = 0$ (cubic anomaly) and $\sum_f Q_f = 0$ (gravitational and mixed anomalies), where the sum runs over all left-handed Weyl fermions.

Gauging $B$ or $L$ separately leads to gauge anomalies that cannot be canceled within the SM fermion content. In contrast, the $U(1)_{B-L}$ combination is anomaly-free when right-handed neutrinos are included, as quarks carry $B-L = 1/3$ while leptons carry $B-L = -1$, leading to natural cancellation generation by generation~\parencite{Marshak:1979fm, Mohapatra:1980yp}. This makes $U(1)_{B-L}$ a particularly well-motivated extension that appears naturally in grand unified theories~\parencite{Fritzsch:1974nn, Georgi:1974sy} and in left-right symmetric models~\parencite{Pati:1974yy, Mohapatra:1974hk, Senjanovic:1975rk}.

An interesting question in BSM model building concerns the possible connection between gauged $B-L$ symmetry and the hypercharge $U(1)_Y$ of the SM. In left-right symmetric models based on the gauge group $SU(3)_C \times SU(2)_L \times SU(2)_R \times U(1)_{B-L}$~\parencite{Pati:1974yy, Mohapatra:1974hk, Senjanovic:1975rk}, the hypercharge emerges from the spontaneous breaking of the extended gauge symmetry. Specifically, the breaking pattern involves two $U(1)$ factors: $U(1)_{B-L}$ and an additional $U(1)$ associated with the right-handed sector. This second factor, denoted $U(1)_{T^3_R}$, corresponds to the diagonal, electrically neutral generator of $SU(2)_R$—the third component of the right-handed weak isospin. The name $T^3_R$ reflects its role as the right-handed analog of the weak isospin component $T^3_L$ in the SM's $SU(2)_L$. In that framework, the SM hypercharge emerges from the breaking pattern
\[
U(1)_{B-L} \times U(1)_{T^3_R} \rightarrow U(1)_Y,
\]
establishing a direct connection between $B-L$ conservation and the electroweak structure. This relationship provides strong theoretical motivation for considering $U(1)_{T^3_R}$ as a gauge extension of the SM~\parencite{Assad:2017iib, MohapatraPati1975, SenjanovicMohapatra1975}. Moreover, it suggests that phenomenological signatures associated with $U(1)_{T^3_R}$ could provide indirect evidence for the underlying left-right symmetric structure and the role of $B-L$ in electroweak symmetry breaking.

In this framework, right-handed SM fermions are charged under $U(1)_{T^3_R}$, leading to distinct phenomenological predictions. In the minimal realization, the $SU(2)_L$ Higgs doublet is a singlet under $U(1)_{B-L}$ and acquires its hypercharge from $U(1)_{T^3_R}$. Its vacuum expectation value (VEV) links the symmetry-breaking scales of $U(1)_Y$ and $U(1)_{T^3_R}$. Alternatively, these scales can be decoupled by introducing an additional $U(1)_G$ group, under which SM fermions are singlets but the Higgs is charged. The inclusion of this symmetry provides model-building freedom to explore a wider range of phenomenological scenarios. For that case, the hypercharge is given by the linear combination
\begin{equation}
Y=Q_{T^3_R}+\frac{1}{2}Q_{B-L} + Q_G.
\end{equation}
More generally, scenarios can be constructed where the hypercharge is not directly related to $U(1)_{T^3_R}$.

In any case, recent theoretical and phenomenological work has focused on models where the low-energy gauge symmetry of the SM is extended by this Abelian group, with the crucial feature that the spontaneous breaking of $U(1)_{T^3_R}$ is not tied to electroweak symmetry breaking. It has been shown to resolve tensions related to lepton flavor violation (LFV) and lepton universality violation (LUV) by explaining the anomalous magnetic moment of the muon $(g-2)_\mu$, $B$ meson anomalies~\parencite{Dutta2022}, and providing viable dark matter candidates~\cite{Dutta2019, Dutta2020b, Dutta:2022qvn}. We are interested in exploring the collider phenomenology of this $U(1)_{T^3_R}$ extension, assuming that the breaking of this symmetry occurs at a scale much higher than the electroweak scale.

The spontaneous breaking of $U(1)_{T^3_R}$ requires the introduction of a complex scalar field $\phi$, which is a singlet under the SM gauge group. When $\phi$ acquires a vacuum expectation value, the symmetry is spontaneously broken, and the associated electrically neutral gauge boson becomes massive by absorbing the CP-odd (pseudoscalar) component of $\phi$ as its longitudinal degree of freedom. The CP-even component of $\phi$ remains in the physical spectrum as a scalar particle, denoted $\phi'$, commonly referred to as the dark Higgs~\parencite{DiLuzio2018, Baker2019, Michaels:2020fzj, Dev:2021otb}. And, to ensure anomaly cancellation in the $U(1)_{T^3_R}$ model, a right-handed neutrino $\nu_R$ is required for each SM generation. In addition, a set of new vector-like fermions $(\chi_\mathrm{u}, \chi_d, \chi_\ell, \chi_\nu)$ is introduced to generate fermion masses in a UV-complete theory, following the universal see-saw mechanism~\parencite{Berezhiani, Chang1987, Davidson1987, Rajpoot1987, Babu1989, Babu1990}, see App.~\ref{app:universal_seesaw}. The universal see-saw mechanism introduces a non-trivial coupling between the top quark and the vector-like quark $\chi_\mathrm{u}$ through the dark Higgs $\phi'$, enabling the vertex $\chi_\mathrm{u} - \mathrm{t} -\phi'$. This coupling allows for the production of $\mathrm{t}\chi_\mathrm{u} \phi'$ final states via $\chi_\mathrm{u}$--$\mathrm{t}$ fusion, see Fig.~\ref{fig:qqfusion}. Since $\chi_\mathrm{u}$ couples to SM quarks and gluons, it can be copiously produced at hadron colliders. Its energetic decay products, together with a $\phi'$ mediator carrying significant transverse momentum, can be efficiently detected, especially if $\phi'$ decays to visible SM particles in the central detector region.

In this chapter, we present a phenomenological study of search strategies at the LHC for a light (GeV-scale) scalar $\phi'$ produced in association with a heavy (TeV-scale) top-partner $\chi_\mathrm{u}$. This work, that  has been published as~\parencite{Qureshi:2024naw}, focuses on the previously unexplored production channel $\mathrm{pp}\to \mathrm{t}\chi_\mathrm{u}\phi'$. This contrasts with more commonly studied processes $\mathrm{pp}\to \chi_\mathrm{u}\chi_\mathrm{u}\to \mathrm{t}\phi'\mathrm{t}\phi'$, and the di-photon $\phi'$ decay channels (see Sec.~\ref{sec:exp} and~\parencite{Bhardwaj_2022, Bhardwaj_2022_2, Bardhan_2023, Banerjee_2016, Alves_2024}).

To further maximize the sensitivity to this complex signal, a central component of our analysis is the use of ML. We employ an analysis based on Boosted Decision Trees (BDT)~\parencite{friedman_greedy_2001}. The BDT output is used in a profile-binned likelihood test to determine the signal significance for each model. The effectiveness of BDTs and other ML algorithms has been demonstrated in numerous experimental and phenomenological studies~\parencite{Ai:2022qvs, ATLAS:2017fak, Biswas:2018snp, Chung:2020ysf, Feng:2021eke, ttZprime, Chigusa:2022svv, Arganda2024, Ajmal_2024, Dutta_2015}. Our results show that the BDT approach significantly improves sensitivity.


The remainder of this chapter is organized as follows. Sec.~\ref{sec:modelU1} describes the minimal $U(1)_{T^3_R}$ model. Sec.~\ref{sec:exp} reviews current relevant LHC results. Sec.~\ref{sec:sims} details the MC simulation samples used in this study. Sec.~\ref{sec:ML} discusses the motivation and implementation of the machine learning workflow, and Sec.~\ref{sec:resultsU1} presents the main results.


\input{Chapters/C03/s02.tex}
\input{Chapters/C03/s01.tex}
\input{Chapters/C03/s03.tex}
\input{Chapters/C03/s04.tex}
\input{Chapters/C03/s05.tex} % Results
\input{Chapters/C03/s06.tex} % Discussion and conclusions