\chapter{Standard Model of Particle Physics}\label{ch:sm}

The Standard Model (SM) of particle physics is a quantum field theory (QFT) that describes matter as fermionic particles and their fundamental interactions. The forces are incorporated through the gauge principle, where force-carrying particles---vector bosons with spin one, arising from the adjoint representation of symmetry groups (\textit{gauge groups})---mediate the interactions between matter particles~\parencite{greiner2000relativistic,pokorski2000gauge}. However, this elegant formulation is not sufficient to account for particle masses. These are generated through Yukawa interactions, which are scalar-fermion couplings between the Higgs field and the fermion fields. While the Yukawa interactions themselves are not gauge interactions, their allowed structure---specifically, which fermions they can couple and their transformation properties---is strictly dictated by the gauge symmetry of the theory. This combined framework of gauge and Yukawa sectors successfully describes three of the four fundamental forces in nature.

In this chapter, we contextualize the SM by introducing the basic concepts of quantum field theory, including the notion of fields and symmetries. We then present the particle content of the SM, its gauge group, and the Lagrangian density that describes its dynamics. The Higgs mechanism and its role in providing mass to the weak gauge bosons and fermions are also discussed. Finally, we address the main deficiencies of the SM and review the experimental evidence that motivates the search for physics beyond the SM.

\input{Chapters/C01/s01.tex} % Fields
\input{Chapters/C01/s02.tex} % Standard Model
\input{Chapters/C01/s03.tex} % Deficiencies of Standard Model and Evidence
\input{Chapters/C01/s04.tex} % Beyond Standard Model